Eleven years of student replication projects provide evidence on the correlates of replicability in psychology

### 2023-04-04 paper/manuscript.Rmd
more intro rambling

@@

@gilbert2016 @anderson2016 people argue a lot about how to interpret RPP

@@

replication is important (but not everything) - good starting place
 * prereq of theory building that the results of interest "can be observed consistently" 
 * but not everything (can still have other problems)

The replication rates in psychology are low, and 
replication rates are low in psychology (see studies)

big open question: why

experimental and correlational approaches to answering it

experimental is hard - protzko, ML5 etc. but they are expensive and you can’t do them for many research questions

correlational has been pursued extensively, but is overfit to small data (prediction markets / ML models / regression of all types)

@@

Psychology is in the middle of heated debate over its research practices and the subject of many reform efforts.  Many of these issues concern how trustworth the published literature is (or isn't), how it got to be this way, and what measures can be taken to improve the quality of findings for the future. These issues center around issues of reproducibility  and replicability. Concerns about replication rates have become a large point of discussion in psychology and other fields, with large scale replications spurring these discussions and providing the empirical data analyzed by all sides.

 A major concern is that published literature may not replicate as much as previously thought or as desired. This has been illustrated by prominent findings/theories not replicating at all (ego depletion/ TMT?). To understand how replicable the literature as a whole is, we need empirical data about replication attempts across a wide range of studies. Most replications are targeted replications of individual results of interest to the researcher, but these are not sampled randomly from the literature, are performed in inconsistent ways, and may be selectively published depending on the results. Thus, they are not ideal for estimating the overall replicability of the field.

There have been a limited number of large-scale replication efforts, which sample studies from a discipline to estimate overall replicability. Due to the arduous process of replicating many studies, there have not been many large-scale replication efforts. Thus, all of the argumentation around predictors of replications and field-wide replication rates is fit to a small number of data points.

## prior literature
We are aware of three large-scale replication efforts replicating experimental results in the existing psychology literature. RP:P sampled roughly 100 studies from top psychology journals in 2008 [@openscienceconsortium2015]. They found an overall replication rate around 40%, which provided evidence to support growing concerns about the non-reliability of the literature. The RP:P dataset was itself much discussed and reanalyzed [@etz2016;@gilbert2016;@patil2016] [idk, maybe worth framing that this was a big deal study when it came out]. TODO some discussion of how many papers have reanalysed this cutting the data different ways.

The ManyLabs series of studies have also done large-scale replications of effects from psychology. Due to their primary goal of investigating different forms of hetereogeneity, their sampling has been non-representive, focusing on short studies with only two conditions. Across Many labs 1-3 foo bar replicated [@klein2014; @klein2018; @ebersole2016]. Many labs 5 was a re-replication attempt on 10 of RP:P and rescued 2/10 [@ebersole2020]. TODO look up details

@camerer2018 replicated the 21 behaviors studies published in Nature and Science from 2010-2015 that did not require special populations or special equipment. They found a roughly 60% replication rate.

@@

can even cite the “science is not a signal detection problem” paper on not randomly sampling from the literature.

Prior approaches to replicability have focused on a potentially problematic estimand: the probability of a finding in the literature being somehow truly replicable. Critics have pointed out that "true" replicability may not be possible to estimate outside of a specific sample (van bavel etc.) or even time period (ramscar).

@@

In contrast, we have pursued a different estimand: the probability that a researcher, on selecting a finding of interest from the literature, can successfully achieve a result satisfactorily close enough to the original that they feel that they can build on it in their own work, with all the necessary compromises to the methods and sample of the original that may be required by the constraints of the situation. SPELL OUT SAMPLING DIFFERENCE AND SUCCESS DIFFERENCE.

>>>>

To understand how replicable the literature as a whole is, we need empirical data about replication attempts across a wide range of studies. Most replications are targeted replications of individual results of interest to the researcher, but these are not sampled randomly from the literature, are performed in inconsistent ways, and may be selectively published depending on the results. Thus, they are not ideal for estimating the overall replicability of the field.

There have been a limited number of large-scale replication efforts, which sample studies from a discipline to estimate overall replicability. Due to the arduous process of replicating many studies, there have not been many large-scale replication efforts. Thus, all of the argumentation around predictors of replications and field-wide replication rates is fit to a small number of data points. 
Psychology is in the middle of heated debate over its research practices and the subject of many reform efforts.  Many of these issues concern how trustworth the published literature is (or isn't), how it got to be this way, and what measures can be taken to improve the quality of findings for the future. These issues center around issues of reproducibility  and replicability. Concerns about replication rates have become a large point of discussion in psychology and other fields, with large scale replications spurring these discussions and providing the empirical data analyzed by all sides.

  [idk, maybe worth framing that this was a big deal study when it came out]. TODO some discussion of how many papers have reanalysed this cutting the data different ways.

 Many labs 5 was a re-replication attempt on 10 of RP:P and rescued 2/10 [@ebersole2020]. TODO look up details

@@

@@

The replication rates in psychology are low, with estimates of replicability of experimental findings hovering somewhere around 50%.   A major concern is that published literature may not replicate as much as previously thought or as desired. This has been illustrated by prominent findings/theories not replicating at all (ego depletion/ TMT? CITATIONS). Some prominent non-replications and concerns about research practices that might impact replication rates led to a few large-scale replication projects, aiming to quantify replication rates across swaths of the literature.

We are aware of three large-scale replication efforts replicating experimental results in the existing psychology literature. RP:P sampled roughly 100 studies from top psychology journals in 2008 [@openscienceconsortium2015]. They found an overall replication rate around 40%, which provided evidence to support growing concerns about the non-reliability of the literature.The ManyLabs series of studies have also done large-scale replications of effects from psychology. Due to their primary goal of investigating different forms of hetereogeneity, their sampling has been non-representive, focusing on short studies with only two conditions. Across Many labs 1-3 foo bar replicated [@klein2014; @klein2018; @ebersole2016]. @camerer2018 replicated the 21 behaviors studies published in Nature and Science from 2010-2015 that did not require special populations or special equipment. They found a roughly 60% replication rate.

These replication studies, and similar ones in other fields, open up big questions about why the replication rates are so low. There are many different theories both about causes and motivations for non-replications (relating to the incentives and research practices of science) and about what studies are most affected by these. There have been both experimental and correlational approaches to understanding the replication rates.

A couple studies have attempted to intervene on the low replication rate, but seeing if certain factors can affect whether studies replicate. For instance, @protzko was a prospective replication study, showing that with a slew of best practices and large samples size (TODO look up details), the replications of FOOBAR novel whatevers had similar effect sizes to the original. Many Labs 5 examined how adding expert advice to a replication process might intervene to improve the replicate rate, and found that it mostly didn't, at least on the 10 studies they looked at TODO CITE. These types of experiments are valuable, but they aren't very scalable because they are expensive and some potential influences on non-replication are not experimentally manipulable. 
Instead, many studies have pursued correlational results about what predicts replication success, whether via regression models, prediction markets and elicitations, or machine learning techniques. However, all of these approaches depend on data from replications. While many replications have been conducted, most were not randomly sampled and are not pooled together, so these studies instead rely on the small number of data points, pulling heavily from RP:P in particular. The RP:P dataset was itself much discussed and reanalyzed [@etz2016;@gilbert2016;@patil2016;@anderson2016].

Our contribution to this literature is to provide a collection of 176 replication projects, done by students for a class. This is a sizeable number of projects, compared to the existing amount of data.

@@

@@

Prior approaches to replicability have focused on a potentially problematic estimand: the probability of a finding in the literature being somehow truly replicable. Critics have pointed out that "true" replicability may not be possible to estimate outside of a specific sample (CITE van bavel etc.) or even time period (CITE ramscar).

@@

In contrast, we have pursued a different estimand: the probability that a researcher, on selecting a finding of interest from the literature, can successfully achieve a result satisfactorily close enough to the original that they feel that they can build on it in their own work, with all the necessary compromises to the methods and sample of the original that may be required by the constraints of the situation.

There's different approaches one could take to sampling, we take the versimilitude approach. Our sample is not a random sample of the literature, or of a given journal, or some weighting, but it does represent studies that people are interested in replicating.

Quantifying when an effect is a replication is hard, and many statistical metrics have been proposed, often with tradeoffs (TODO citations). We instead focus on a subjective score representing a more functionally oriented outcome: studies replicate if the results match the original enough that researchers would feel comfortable building on the study.  This matches the motivation for many replications (often in replicate and extend paradigms). SPELL OUT SAMPLING DIFFERENCE AND SUCCESS DIFFERENCE.

There's a "can I trust this finding" question, but a key part of it is, "if I did this, would I get the same pattern of results" -- and that's what we're measuring.

### 2023-04-04 paper/manuscript.tex
more intro rambling

circle back to the idea that student replications (either as class or before extension into their own work) would be valuable
\# Citation slush pile

@@

Anderson, C., Bahnik, S., Barnett-Cowan, M., Bosco, F., Chandler, J., Chartier, C., \& otherss. (2016). Response to comment on ``estimating the reproducibility of psychological science.'' Science, 351(6277), 1037--1037.

Gilbert, D. T., King, G., Pettigrew, S., \& Wilson, T. D. (2016). Comment on ``estimating the reproducibility of psychological science.'' Science, 351(6277), 1037--1037.

Schmidt, S. (2009). Shall we really do it again? The powerful concept of replication is neglected in the social sciences. Review of General Psychology, 13, 90--100.

Simonsohn, U. (2015). Small telescopes: Detectability and the evaluation of replication results. Psychol. Sci., 26(5), 559--569.

Gelman, A. (2018). Don't characterize replications as successes or failures. Behavioral and Brain Sciences, 41.

Mathur, M. B., \& VanderWeele, T. J. (2020a). New statistical metrics for multisite replication projects. J. R. Stat. Soc. Ser. A Stat. Soc., 183(3), 1145--1166.

pedagogy for replication in intro: frank \& saxe, hawkins et al., quintana nature paper, scarcity PNAS pedagogy paper,

Collaborative replications and education project (CREP)
JE Grahe, MJ Brandt, H IJzerman, N Legate, J Wagge, YJ Weisberg, BJ Wiggins, Retrieved from osf. io/wfc6u, 2016 - Cited by 29

\url{https://psycnet.apa.org/record/2022-15361-001}
\url{https://journals.sagepub.com/doi/pdf/10.1177/1475725719868149}
\url{https://content.iospress.com/articles/education-for-information/efi190308}

\url{https://psycnet.apa.org/fulltext/2018-11647-005.html}

@@

@@

@@

@@

@@

@@

@@

@@

\hypertarget{appendix-appendices}{%
\appendix}

\hypertarget{appendix-a}{%
\section*{Appendix A}\label{appendix-a}}
\addcontentsline{toc}{section}{Appendix A}

Some appendix text.

\hypertarget{appendix-b}{%
\section*{Appendix B}\label{appendix-b}}
\addcontentsline{toc}{section}{Appendix B}

More appendix text.

>>>>

circle back to the idea that student replications (either as class or before extension into their own work) would be valuable

\hypertarget{citation-slush-pile}{%
\section{Citation slush pile}\label{citation-slush-pile}}

@@

Gilbert et al. (\protect\hyperlink{ref-gilbert2016}{2016}) Anderson et al. (\protect\hyperlink{ref-anderson2016}{2016}) people argue a lot about how to interpret RPP

Schmidt (\protect\hyperlink{ref-schmidt2009}{2009}) early theory of replication as part of advancing knowledge

Simonsohn (\protect\hyperlink{ref-simonsohn2015}{2015}) small telescopes

Gelman (\protect\hyperlink{ref-gelman2018}{2018/ed}) Don't characterize replications as successes or failures.

Mathur \& VanderWeele (\protect\hyperlink{ref-mathur2020}{2020}) other metrics of replicability

pedagogy for replication in intro:

Frank \& Saxe (\protect\hyperlink{ref-frank2012}{2012})
Hawkins et al. (\protect\hyperlink{ref-hawkins}{n.d.})

O'Donnell et al. (\protect\hyperlink{ref-odonnell2021}{2021}) reps of some set of related things

Wagge et al. (\protect\hyperlink{ref-wagge2019}{2019}) get UGs to do replications for

Quintana (\protect\hyperlink{ref-quintana2021}{2021}) do reps as honors thesis
Pownall et al. (\protect\hyperlink{ref-pownall2021}{2021}) sips 2021 on teaching openness
Jekel et al. (\protect\hyperlink{ref-jekel2020}{2020}) teach open science w/ replication
Jern (\protect\hyperlink{ref-jern2018}{2018}) in one study didn't show critical thinking benefits for students to make them do replications

@@

Replicability is a desirable feature for any study that we build theory or future experiments off of. It's a pre-requisite for building theories that we have phenomena for these studies to explain, but results that cannot be observed consistently aren't results to build theories around. However, to be a load-bearing empirical result, a study needs to do much more than just replicate, it also needs to be interesting, have a believable theoretical interpretation, generalize appropriately, and have reliable constructs.

@@

\leavevmode\vadjust pre{\hypertarget{ref-anderson2016}{}}%
Anderson CJ, Bahník Š, Barnett-Cowan M, Bosco FA, Chandler J, Chartier CR, Cheung F, Christopherson CD, Cordes A, Cremata EJ, Della Penna N, Estel V, Fedor A, Fitneva SA, Frank MC, Grange JA, Hartshorne JK, Hasselman F, Henninger F, Hulst M van der, Jonas KJ, Lai CK, Levitan CA, Miller JK, Moore KS, Meixner JM, Munafò MR, Neijenhuijs KI, Nilsonne G, Nosek BA, Plessow F, Prenoveau JM, Ricker AA, Schmidt K, Spies JR, Stieger S, Strohminger N, Sullivan GB, Aert RCM van, Assen MALM van, Vanpaemel W, Vianello M, Voracek M, Zuni K (2016) Response to {Comment} on {``{Estimating} the reproducibility of psychological science.''} \emph{Science} \textbf{351}:1037--1037. doi:\href{https://doi.org/10.1126/science.aad9163}{10.1126/science.aad9163}

@@

\leavevmode\vadjust pre{\hypertarget{ref-gelman2018}{}}%
Gelman A (2018/ed) Don't characterize replications as successes or failures. \emph{Behav Brain Sci} \textbf{41}:e128. doi:\href{https://doi.org/10.1017/S0140525X18000638}{10.1017/S0140525X18000638}

@@

\leavevmode\vadjust pre{\hypertarget{ref-jekel2020}{}}%
Jekel M, Fiedler S, Allstadt Torras R, Mischkowski D, Dorrough AR, Glöckner A (2020) How to {Teach Open Science Principles} in the {Undergraduate Curriculum}---{The Hagen Cumulative Science Project}. \emph{Psychol Learn Teach} \textbf{19}:91--106. doi:\href{https://doi.org/10.1177/1475725719868149}{10.1177/1475725719868149}

\leavevmode\vadjust pre{\hypertarget{ref-jern2018}{}}%
Jern A (2018) A preliminary study of the educational benefits of conducting replications in the classroom. \emph{Scholarsh Teach Learn Psychol} \textbf{4}:64--68. doi:\href{https://doi.org/10.1037/stl0000104}{10.1037/stl0000104}

@@

\leavevmode\vadjust pre{\hypertarget{ref-mathur2020}{}}%
Mathur MB, VanderWeele TJ (2020) New statistical metrics for multisite replication projects. \emph{J R Stat Soc Ser A Stat Soc} \textbf{183}:1145--1166. doi:\href{https://doi.org/10.1111/rssa.12572}{10.1111/rssa.12572}

@@

\leavevmode\vadjust pre{\hypertarget{ref-odonnell2021}{}}%
O'Donnell M, Dev AS, Antonoplis S, Baum SM, Benedetti AH, Brown ND, Carrillo B, Choi AL, Connor P, Donnelly K, Ellwood-Lowe ME, Foushee R, Jansen R, Jarvis SN, Lundell-Creagh R, Ocampo JM, Okafor GN, Azad ZR, Rosenblum M, Schatz D, Stein DH, Wang Y, Moore DA, Nelson LD (2021) Empirical audit and review and an assessment of evidentiary value in research on the psychological consequences of scarcity. \emph{Proc Natl Acad Sci USA} \textbf{118}:e2103313118. doi:\href{https://doi.org/10.1073/pnas.2103313118}{10.1073/pnas.2103313118}

@@

\leavevmode\vadjust pre{\hypertarget{ref-pownall2021}{}}%
Pownall M, Azevedo F, Aldoh A, Elsherif M, Vasilev M, Pennington CR, Robertson O, Tromp MV, Liu M, Makel MC, Tonge N, Moreau D, Horry R, Shaw J, Tzavella L, McGarrigle R, Talbot C, Parsons S (2021) Embedding open and reproducible science into teaching: {A} bank of lesson plans and resources. \emph{Scholarsh Teach Learn Psychol}:No Pagination Specified--No Pagination Specified. doi:\href{https://doi.org/10.1037/stl0000307}{10.1037/stl0000307}

\leavevmode\vadjust pre{\hypertarget{ref-quintana2021}{}}%
Quintana DS (2021) Replication studies for undergraduate theses to improve science and education. \emph{Nat Hum Behav} \textbf{5}:1117--1118. doi:\href{https://doi.org/10.1038/s41562-021-01192-8}{10.1038/s41562-021-01192-8}

\leavevmode\vadjust pre{\hypertarget{ref-schmidt2009}{}}%
Schmidt S (2009) Shall we {Really} do it {Again}? {The Powerful Concept} of {Replication} is {Neglected} in the {Social Sciences}. \emph{Rev Gen Psychol} \textbf{13}:90--100. doi:\href{https://doi.org/10.1037/a0015108}{10.1037/a0015108}

\leavevmode\vadjust pre{\hypertarget{ref-simonsohn2015}{}}%
Simonsohn U (2015) Small {Telescopes}: {Detectability} and the {Evaluation} of {Replication Results}. \emph{Psychol Sci} \textbf{26}:559--569. doi:\href{https://doi.org/10.1177/0956797614567341}{10.1177/0956797614567341}

\leavevmode\vadjust pre{\hypertarget{ref-wagge2019}{}}%
Wagge JR, Brandt MJ, Lazarevic LB, Legate N, Christopherson C, Wiggins B, Grahe JE (2019) \href{https://www.frontiersin.org/articles/10.3389/fpsyg.2019.00247}{Publishing {Research With Undergraduate Students} via {Replication Work}: {The Collaborative Replications} and {Education Project}}. \emph{Front Psychol} \textbf{10}

@@
